XXXXX

### 2014-12-16 paper/RDF validation requirements/RDF validation requirements.tex
RDF validation requirements tex file has been updated.

\begin{itemize}
	\item DL termonology: universal restriction
\end{itemize}

\tb{definition}

@@

\tb{definition}

@@

\tb{definition}

@@

@@

\begin{ex}
# Individuals matching the ’:SuperJediMaster’ data shape
# ------------------------------------------------------
:Yoda 
\end{ex}

\tb{ToDO: add DL}

@@

\begin{itemize}
	\item A minimum cardinality expression ObjectMinCardinality( n OPE CE ) consists of a nonnegative integer n, an object property expression OPE, and a class expression CE, and it contains all those individuals that are connected by OPE to at least n different individuals that are instances of CE. If CE is missing, it is taken to be owl:Thing.
	\item A minimum cardinality expression DataMinCardinality( n DPE DR ) consists of a nonnegative integer n, a data property expression DPE, and a unary data range DR, and it contains all those individuals that are connected by DPE to at least n different literals in DR. If DR is not present, it is taken to be rdfs:Literal. 
	\item qualified: CE respective DR is stated.
\end{itemize}

@@

\tb{ToDO: check DL}

@@

\tb{ToDO: check DL}

\begin{gcotable}
property & foaf:Person & hasName & - & xsd:string & $\geq$ & 2 \\
\end{gcotable}

@@

\tb{ToDO: OWL 2 DL}

@@

\tb{ToDO: add DL}

@@

\tb{ToDO: check DL}

@@

@@

@@

\tb{ToDO: add DL}

@@

\begin{itemize}
	\item A maximum cardinality expression ObjectMaxCardinality( n OPE CE ) consists of a nonnegative integer n, an object property expression OPE, and a class expression CE, and it contains all those individuals that are connected by OPE to at most n different individuals that are instances of CE. If CE is missing, it is taken to be owl:Thing.
	\item A maximum cardinality expression DataMaxCardinality( n DPE DR ) consists of a nonnegative integer n, a data property expression DPE, and a unary data range DR, and it contains all those individuals that are connected by DPE to at most n different literals in DR. If DR is not present, it is taken to be rdfs:Literal. 
	\item qualified: CE respective DR is stated. 
\end{itemize}

@@

\tb{ToDO: OWL 2 DL}

@@

\tb{ToDO: add DL}

@@

@@

\tb{ToDO: add DL}

\section{Exact Qualified Cardinality Restrictions on Properties}

@@

\tb{ToDO: OWL 2 DL}

@@

\tb{ToDO: add DL}

@@

\begin{DL}
marriedTo $\equiv$ $marriedTo^{-}$ \\
\end{DL}

@@

\begin{DL}
Disjoint( parentOf , $parentOf^{-}$ )
\end{DL}

@@

\begin{DL}
$\top$ $\sqsubseteq$ $\neg$ $\exists$ marriedTo . Self
\end{DL}

@@

\begin{DL}
$A \equiv \forall foaf:name . xsd:string$ \\
$B \equiv \forall foaf:givenName . xsd:string$ \\
$C \equiv A XOR B$ \\
$D \equiv \forall foaf:mbox . xsd:string$ \\
$Human \equiv C \sqcap D$ \\
\end{DL}

\section{Context-Specific Exclusive OR of Properties}

@@

@@

@@

@@

\begin{DL}
Jedi $\equiv$ $\exists$ laserSwordColor . \{'blue'\} $\sqcup$ \{'green'\} $\sqcup$ \{'white'\} \\
\end{DL}

\tb{ToDo: check DL}

@@

\begin{DL}
Sith $\equiv$ $\neg$( $\exists$ laserSwordColor . \{'blue'\} $\sqcup$ \{'green'\} $\sqcup$ \{'white'\} ) \\
\end{DL}

\tb{ToDo: check DL}

@@

\begin{itemize}
	\item \ms{skos:Concept} resources can only be members of the listed controlled vocabularies
\end{itemize}

\begin{DL}
$A \equiv skos:Concept \sqcap B$ \\
$B \equiv \forall skos:inScheme . C$ \\
$C \equiv skos:ConceptScheme \sqcap ( \{:BookSubjects\} \sqcup \{:BookTopics\} \sqcup \{:BookCategories\} )$ \\
\end{DL}

>>>>

\begin{itemize}
	\item DL terminology: value restriction
\end{itemize}

$(\forall R.C)^\mathcal{I}= \{a \in \Delta^\mathcal{I}\mid (a,b) \in R^\mathcal{I} \rightarrow b \in C^\mathcal{I} \}$ where $\cdot^\mathcal{I}$ is an interpretation function, $\Delta^I$ is the domain, $a,b \in \Delta^\mathcal{I}$ are individuals $C$ is a concept, $R$ is a role.

@@

$\exists R. \top \sqsubseteq C$ is the object property restriction where $R$ is the object property (role) whose domain is restricted to concept $C$.

@@

$\top \sqsubseteq \forall R . C$ is the range restriction to object property $R$ (restricted by concept $C$).

@@

Definition:\\

$\leq n R. \top$ is minimum unqualified cardinality restriction where $n \in \mathbb{N}$ (written $\leq  n R$ in short).

@@

\begin{ex}
# Individuals matching the ’:SuperJediMaster’ data shape
# ------------------------------------------------------
:Yoda 
\end{ex}
\er{Note that this example includes also has at-most restriction.}\\

Assume that  mentorOf and studentOf are inverse properties.\\
\begin{DL}
$Jedi \sqsubseteq \exists attitude.\{good\} $\\
$JediStudent \sqsubseteq Jedi \sqcap 1\geq studentOf.\top \sqcap 1\leq studentOf.\top$ \\
$JediMasters \sqsubseteq Jedi \sqcap 1\leq studentOf^{-}.\top \sqcap 2\geq studentOf^{-}.\top $\\
$SuperJediMaster \sqsubseteq Jedi \sqcap  3\leq studentOf^{-}.\top $
\end{DL}

@@

\begin{itemize}
	\item A minimum cardinality expression ObjectMinCardinality( n OPE CE ) consists of a nonnegative integer n, an object property expression OPE, and a class expression CE, and it contains all those individuals that are connected by OPE to at least n different individuals that are instances of CE. If CE is missing, it is taken to be owl:Thing.
	\item A minimum cardinality expression DataMinCardinality( n DPE DR ) consists of a nonnegative integer n, a data property expression DPE, and a unary data range DR, and it contains all those individuals that are connected by DPE to at least n different literals in DR. If DR is not present, it is taken to be rdfs:Literal. 
	\item qualified: CE respective DR is stated.
\end{itemize}
Definition:\\

$\leq n R. C$ is minimum qualified cardinality restriction where $n \in \mathbb{N}$.

@@

@@

\tb{ToDO: check DL}
\er{foaf: or xsd:string is no DL syntax. So you can delete 18.2 if you are fine.}
\begin{gcotable}
property & foaf:Person & hasName & - & xsd:string & $\geq$ & 2 \\
\end{gcotable}

@@

%\tb{ToDO: OWL 2 DL}

@@

\noindent $mentorOf$ and $studentOf$ are taken to be inverse properties:\\

\begin{DL}
$Jedi \sqsubseteq \exists attitude.\{good\} $\\
$JediStudent \sqsubseteq Jedi \sqcap 1\geq studentOf.Jedi \sqcap 1\leq studentOf.Jedi$ \\
$JediMasters \sqsubseteq Jedi \sqcap 1\leq studentOf^{-}.Jedi\sqcap 2\geq studentOf^{-}.Jedi $\\
$SuperJediMaster \sqsubseteq Jedi \sqcap  3\leq studentOf^{-}.Jedi $
\end{DL}
\subsection{Complex Example}

@@

%\tb{ToDO: check DL}

@@

Definition:\\

$\geq n R. \top$ is maximum unqualified cardinality restriction where $n \in \mathbb{N}$ (written $\geq  n R$ in short).
\subsection{Simple Example}

@@

\begin{itemize}
	\item Jedis have the attitude 'good'
	\item Jedi students are students of exactly 1 resource
	\item Jedi masters are mentoring at least 1 and at most 2 resources
	\item Super Jedi masters are mentoring at least 3 resources  
\end{itemize}
\er{Notice that this example includes minimum restriction.}
\begin{ex}
# data
# ----
:Yoda 
    :attitude 'good' ;
    :mentorOf :MaceWindu , :Obi-Wan , :Luke .
:MaceWindu
    :attitude 'good' ;
    :studentOf :Yoda .
:Obi-Wan 
    :attitude 'good' ;
    :studentOf :Yoda ;
    :mentorOf :Anakin .
:Anakin
    :attitude 'good' ; 
    :studentOf :Obi-Wan .
:Luke
    :attitude 'good' ;
    :studentOf :Yoda .
\end{ex}

@@

%\tb{ToDO: add DL}
\noindent Taking mentorOf and studentOf as inverse properties. \\
\begin{DL}
$Jedi \sqsubseteq \exists attitude.\{good\} $\\
$JediStudent \sqsubseteq Jedi \sqcap 1\geq studentOf.\top \sqcap 1\leq studentOf.\top$ \\
$JediMasters \sqsubseteq Jedi \sqcap 1\leq studentOf^{-}.\top \sqcap 2\geq studentOf^{-}.\top $\\
$SuperJediMaster \sqsubseteq Jedi \sqcap  3\leq studentOf^{-}.\top $
\end{DL}

@@

\begin{itemize}
	\item A maximum cardinality expression ObjectMaxCardinality( n OPE CE ) consists of a nonnegative integer n, an object property expression OPE, and a class expression CE, and it contains all those individuals that are connected by OPE to at most n different individuals that are instances of CE. If CE is missing, it is taken to be owl:Thing.
	\item A maximum cardinality expression DataMaxCardinality( n DPE DR ) consists of a nonnegative integer n, a data property expression DPE, and a unary data range DR, and it contains all those individuals that are connected by DPE to at most n different literals in DR. If DR is not present, it is taken to be rdfs:Literal. 
	\item qualified: CE respective DR is stated. 
\end{itemize}
Definition:\\

$\leq n R. C$ is maximum qualified cardinality restriction where $n \in \mathbb{N}$.

@@

\tb{ToDO: OWL 2 DL}
\er{why OWL 2 DL?}
\begin{ex}
# ShEx
# ----
:Jedi {
    :attitude ('good') }
:JediStudent {
    & :Jedi ,
    :studentOf @:Jedi{1} }
:JediMaster {
    & :Jedi ,
    :mentorOf @:Jedi{1,2} }
:SuperJediMaster {
    & :Jedi ,
    :mentorOf @:Jedi{3,} }
\end{ex}

@@

%\tb{ToDO: add DL}

\noindent $mentorOf$ and $studentOf$ are taken to be inverse properties:\\

\begin{DL}
$Jedi \sqsubseteq \exists attitude.\{good\} $\\
$JediStudent \sqsubseteq Jedi \sqcap 1\geq studentOf.Jedi \sqcap 1\leq studentOf.Jedi$ \\
$JediMasters \sqsubseteq Jedi \sqcap 1\leq studentOf^{-}.Jedi\sqcap 2\geq studentOf^{-}.Jedi $\\
$SuperJediMaster \sqsubseteq Jedi \sqcap  3\leq studentOf^{-}.Jedi $
\end{DL}

@@

Definition:\\

$\geq n R. \top \sqcap \leq n R. \top $ is exact unqualified cardinality restriction where $n \in \mathbb{N}$. ($\top$ can be dropped in short writing.)
\subsection{Simple Example}

@@

%\tb{ToDO: add DL}
\begin{DL}
$Jedi \sqsubseteq \exists attitude.\{good\} $\\
$JediStudent \sqsubseteq Jedi \sqcap 1\geq studentOf.\top \sqcap 1\leq studentOf.\top$ \\
\end{DL}
\section{Exact Qualified Cardinality Restrictions on Properties}

@@

Definition:\\

$\geq n R. C \sqcap \leq n R. C $ is exact unqualified cardinality restriction where $n \in \mathbb{N}$.
\subsection{Simple Example}

%\tb{ToDO: OWL 2 DL}

@@

\begin{DL}
$Jedi \sqsubseteq \exists attitude.\{good\} $\\
$JediStudent \sqsubseteq Jedi \sqcap 1\geq studentOf.Jedi \sqcap 1\leq studentOf.Jedi$ \\
\end{DL}

@@

\begin{DL}
$marriedTo \sqsubseteq marriedTo^{-}$ \\
\end{DL}

@@

\begin{DL}
$parentOf \sqsubseteq \neg parentOf^{-}$ or\\
$parentOf \sqcap parentOf^{-} \sqsubseteq \bot$
\end{DL}

@@

\begin{DL}
$\top$ $\sqsubseteq$ $\neg$ $\exists  marriedTo . Self$   OR\\
$\exists marriedTo . Self \sqsubseteq \bot$  OR\\
$marriedTo \sqcap marriedTo^{-} \sqsubseteq \bot$ (without special self-concept Self)
\end{DL}

@@

\begin{DL}
$A \equiv \forall foaf:name . xsd:string$ \\
$B \equiv \forall foaf:givenName . xsd:string$ \\
$C \equiv A XOR B$ \\
$D \equiv \forall foaf:mbox . xsd:string$ \\
$Human \equiv C \sqcap D$ \\
\end{DL}
\er{Where does this using $foaf:$ in DL come from? We need a description.}
\section{Context-Specific Exclusive OR of Properties}

@@

\er{Let us mark that $foaf:$ occurs here.}

@@

\er{Unsure what is meant in this example.}

@@

%\tb{ToDo: DL}

\er{Unsure what is meant in this example. Could be instead:}

\begin{DL}
$Person \sqsubseteq ((Male \sqcap \neg Female) \sqcup (\neg Male \sqcap  Female)) $ 
\end{DL}

@@

\begin{DL}
Jedi $\equiv$ $\exists$ laserSwordColor . \{blue, green, white\} \\
\end{DL}

%\tb{ToDo: check DL}

@@

\begin{DL}
Sith $\equiv$ $\neg$ $\exists$ laserSwordColor . \{blue, green, white\} \\
\end{DL}

%\tb{ToDo: check DL}

@@

\begin{itemize}
	\item \ms{skos:Concept} resources can only be members of the listed controlled vocabularies
\end{itemize}
\er{This sort of DL-syntax (with semicolon) must be rechecked. Note that in DL, it does not have to correspond directly to names of things in machine-readable syntax. It is fine to say BookSubjects rather than :BookSubjects for the sake of simplicity }
\begin{DL}
$A \equiv skos:Concept \sqcap B$ \\
$B \equiv \forall skos:inScheme . C$ \\
$C \equiv skos:ConceptScheme \sqcap ( \{:BookSubjects\} \sqcup \{:BookTopics\} \sqcup \{:BookCategories\} )$ \\
\end{DL}
